---
title: "DATA3888 Optiver Project"
output: html_document
date: "2024-05-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load necessary libraries
library(e1071)
library(tidyverse)
library(randomForest)
library(pracma)
library(miscTools)
```

```{r, message=FALSE}
#stock = read.csv('Optiver_additional data/order_book_feature.csv', sep='\t')
#save(stock, file='stock.RData')
#target = read.csv('Optiver_additional data/order_book_target.csv', sep='\t')
#save(target, file='target.RData')

load('../data/stock.RData')
load('../data/target.RData')
stock_IDs = unique(stock['stock_id'])

# Group timeIDs by quarters
time_IDs = read.csv('../data/time_id_reference.csv')
Q1 = time_IDs |> filter(date >= '2021-01-01' & date <= '2021-03-31') 
Q2 = time_IDs |> filter(date >= '2021-04-01' & date <= '2021-06-30')
Q3 = time_IDs |> filter(date >= '2021-07-01' & date <= '2021-09-30')
Q4 = time_IDs |> filter(date >= '2021-10-01' & date <= '2021-12-31')
```

```{r}
# Select a random date in each quarter
set.seed(1)
Q1_date = sample(unlist(unique(Q1['date'])),1)
Q2_date = sample(unlist(unique(Q2['date'])),1)
Q3_date = sample(unlist(unique(Q3['date'])),1)
Q4_date = sample(unlist(unique(Q4['date'])),1)

# Get time IDs in selected dates
Q1_timeIDs = time_IDs |> filter(date == Q1_date) |> pull(time_id)
Q2_timeIDs = time_IDs |> filter(date == Q2_date) |> pull(time_id)
Q3_timeIDs = time_IDs |> filter(date == Q3_date) |> pull(time_id)
Q4_timeIDs = time_IDs |> filter(date == Q4_date) |> pull(time_id)
quarters = list(Q1_timeIDs, Q2_timeIDs, Q3_timeIDs, Q4_timeIDs)
```

## Feature engineering

```{r}
# Function to compute realised volatility
comp_vol <- function(x) {
    return(sqrt(sum(x ^ 2)))
}
# Function to compute realised quarticity
comp_quar <- function(x) {
    return(length(x) / 3 * sum(x ^ 4))
}

# Compute volatility and features for each quarter
q = 1
quarters.vol = quarters.features = quarters.quarticity = list()

for(Q_timeIDs in quarters){
  Q_feature = stock |> 
  filter(stock_id == 8382 & time_id %in% Q_timeIDs) |> 
  mutate(BidAskSpread = (ask_price1 / bid_price1) - 1,
         WAP = (bid_price1*ask_size1 + ask_price1*bid_size1)/(bid_size1 + ask_size1),
         num_order = bid_size1 + ask_size1 + bid_size2 + ask_size2,
         volume_imbalance = abs((ask_size1 + ask_size2)-(bid_size1 + bid_size2)),
         bid_spread = bid_price1 - bid_price2,
         ask_spread = ask_price1 - ask_price2,
         time_bucket = ceiling(seconds_in_bucket / 600),)

  Q_target = target |> 
    filter(stock_id == 8382 & time_id %in% Q_timeIDs) |> 
    mutate(BidAskSpread = (ask_price1 / bid_price1) - 1,
           WAP = (bid_price1*ask_size1 + ask_price1*bid_size1)/(bid_size1 + ask_size1),
           num_order = bid_size1 + ask_size1 + bid_size2 + ask_size2,
           volume_imbalance = abs((ask_size1 + ask_size2)-(bid_size1 + bid_size2)),
           bid_spread = bid_price1 - bid_price2,
           ask_spread = ask_price1 - ask_price2,
           time_bucket = ceiling(seconds_in_bucket / 600))
  
  # Compute log-returns
  log_r1 = list()
  for (i in 1 : length(Q_timeIDs)) {
    sec1 = Q_feature |> filter(time_id == Q_timeIDs[i]) |> pull(seconds_in_bucket)
    price1 = Q_feature |> filter(time_id == Q_timeIDs[i]) |> pull(WAP)
    
    sec2 = Q_target |>  filter(time_id == Q_timeIDs[i]) |>  pull(seconds_in_bucket)
    price2 = Q_target |>  filter(time_id == Q_timeIDs[i]) |> pull(WAP)
    
    sec = c(sec1, sec2)
    price = c(price1, price2)
    
    log_r <- log(price[-1] / price[1:(length(price) - 1)])
    log_r1[[i]] <- data.frame(time = sec[-1], log_return = log_r)
    time.no.change <- (1:3599)[!(1:3599 %in% log_r1[[i]]$time)]
    if (length(time.no.change) > 0) {
      new.df <- data.frame(time = time.no.change, log_return = 0)
      log_r1[[i]] <- rbind(log_r1[[i]], new.df)
      log_r1[[i]] <- log_r1[[i]][order(log_r1[[i]]$time), ]
    }
  }

  # Compute volatility
  vol = list()
  for (i in 1 : length(log_r1)) {
    log_r1[[i]] <- log_r1[[i]] %>% mutate(time_bucket = ceiling(time / 600))
    vol[[i]] <- aggregate(log_return ~ time_bucket, data = log_r1[[i]], FUN = comp_vol)
    colnames(vol[[i]]) <- c('time_bucket', 'volatility')
  }
  quarters.vol[[q]] = do.call(rbind, vol)
  
  # Compute features
  mean.price.df = var.price.df = mean.order.df = var.order.df = mean.BAS.df = var.BAS.df = 
    mean.volIm.df = var.volIm.df = mean.bidSpread.df = var.bidSpread.df = 
    mean.askSpread.df = var.askSpread.df = data.frame()
  
  for (i in 1 : length(vol)) {
    stats.bucket1 = Q_feature %>% 
      filter(time_id == Q_timeIDs[i] & time_bucket != 0) %>% 
      select(c(BidAskSpread,WAP,num_order,volume_imbalance,bid_spread,ask_spread, time_bucket)) 
    
    stats.bucket2 = Q_target %>% 
      filter(time_id == Q_timeIDs[i]) %>% 
      select(c(BidAskSpread,WAP,num_order,volume_imbalance,bid_spread,ask_spread, time_bucket))
    
    stats.bucket = rbind(stats.bucket1, stats.bucket2)
    
    # For each 10-min time bucket, we compute the following statistics
    mean.price <- aggregate(WAP ~ time_bucket, data = stats.bucket, FUN = mean)
    var.price <- aggregate(WAP ~ time_bucket, data = stats.bucket, FUN = var)
    mean.order <- aggregate(num_order ~ time_bucket, data = stats.bucket, FUN = mean)
    var.order <- aggregate(num_order ~ time_bucket, data = stats.bucket, FUN = var)
    mean.BAS <- aggregate(BidAskSpread ~ time_bucket, data = stats.bucket, FUN = mean)
    var.BAS <- aggregate(BidAskSpread ~ time_bucket, data = stats.bucket, FUN = var)
    mean.volIm <- aggregate(volume_imbalance ~ time_bucket, data = stats.bucket, FUN = mean)
    var.volIm <- aggregate(volume_imbalance ~ time_bucket, data = stats.bucket, FUN = var)
    mean.bidSpread <- aggregate(bid_spread ~ time_bucket, data = stats.bucket, FUN = mean)
    var.bidSpread <- aggregate(bid_spread ~ time_bucket, data = stats.bucket, FUN = var)
    mean.askSpread <- aggregate(ask_spread ~ time_bucket, data = stats.bucket, FUN = mean)
    var.askSpread <- aggregate(ask_spread ~ time_bucket, data = stats.bucket, FUN = var)
    
    mean.price.df = rbind(mean.price.df, mean.price)
    var.price.df = rbind(var.price.df, var.price)
    mean.order.df = rbind(mean.order.df, mean.order)
    var.order.df = rbind(var.order.df, var.order)
    mean.BAS.df = rbind(mean.BAS.df, mean.BAS)
    var.BAS.df = rbind(var.BAS.df, var.BAS)
    mean.volIm.df = rbind(mean.volIm.df, mean.volIm)
    var.volIm.df = rbind(var.volIm.df, var.volIm)
    mean.bidSpread.df = rbind(mean.bidSpread.df, mean.bidSpread)
    var.bidSpread.df = rbind(var.bidSpread.df, var.bidSpread)
    mean.askSpread.df = rbind(mean.askSpread.df, mean.askSpread)
    var.askSpread.df = rbind(var.askSpread.df, var.askSpread)
  }
  quarters.features[[q]] = data.frame(time_bucket = mean.price.df$time_bucket,
                                       mean_price = mean.price.df$WAP,
                                       var_price = var.price.df$WAP,
                                       mean_order = mean.order.df$num_order,
                                       var_order = var.order.df$num_order,
                                       mean_BAS = mean.BAS.df$BidAskSpread,
                                       var_BAS = var.BAS.df$BidAskSpread,
                                       mean_volIm = mean.volIm.df$volume_imbalance, 
                                       var_volIm = var.volIm.df$volume_imbalance,
                                       mean_bidSpread = mean.bidSpread.df$bid_spread, 
                                       var_bidSpread = var.bidSpread.df$bid_spread,
                                       mean_askSpread = mean.askSpread.df$ask_spread, 
                                       var_askSpread = var.askSpread.df$ask_spread)
  
  # Compute quarticity 
  quar = list()
  for (i in 1 : length(log_r1)) {
    quar[[i]] <- aggregate(log_return ~ time_bucket, data = log_r1[[i]], FUN = comp_quar)
    colnames(quar[[i]]) <- c('time_bucket', 'quarticity')
  }
  quarters.quarticity[[q]] = do.call(rbind, quar)
  
  q = q+1
}
```

### Backward selection

We perform backward selection on each quarter, and select features that are consistently significant across the four quarters.

```{r}
reg = quarters.features[[1]][1:35,]
reg$volatility = quarters.vol[[1]][2:36,2]
lm1 = lm(volatility ~ ., reg, weights = 0.8 ^ ((35 - 1):0))
step.back.aic = step(lm1, direction="backward", trace=FALSE)
summary(step.back.aic)
```

```{r}
reg = quarters.features[[2]][1:35,]
reg$volatility = quarters.vol[[2]][2:36,2]
lm1 = lm(volatility ~ ., reg, weights = 0.8 ^ ((35 - 1):0))
step.back.aic = step(lm1, direction="backward", trace=FALSE)
summary(step.back.aic)
```

```{r}
reg = quarters.features[[3]][1:35,]
reg$volatility = quarters.vol[[3]][2:36,2]
lm1 = lm(volatility ~ ., reg, weights = 0.8 ^ ((35 - 1):0))
step.back.aic = step(lm1, direction="backward", trace=FALSE)
summary(step.back.aic)
```

```{r}
reg = quarters.features[[4]][1:35,]
reg$volatility = quarters.vol[[4]][2:36,2]
lm1 = lm(volatility ~ ., reg, weights = 0.8 ^ ((35 - 1):0))
step.back.aic = step(lm1, direction="backward", trace=FALSE)
summary(step.back.aic)
```

## Rolling window forecasting

### Linear regression

```{r}
fit_lm = function(Q_vol, Q_features){
  # Rolling window
  window_size = 12 # 2 hours = 12 buckets of 10-min
  window_lb = 2 # start from 2nd time bucket
  increment = 1 # increment by 10 minutes each time
  max_index = nrow(Q_vol)
  i = 1
  
  lm.models = pred.lm = list()
  r.squared.lm = MSE.lm = QLIKE.lm = MAE.lm = RMSPE.lm = vector()
  
  while(max_index > window_lb + window_size - 1){
    window_ub = window_lb + (window_size - 1)
    # Use 2-hour data to train model (12 10-min time buckets)
    Q.vol.train = Q_vol[window_lb:window_ub,]
    # Actual volatility of the next 10 minutes (next time bucket)
    Q.vol.val = Q_vol[window_ub + 1,]
    
    idx = (window_lb - 1):(window_ub - 1)
    reg = Q_features[idx,]
    reg$volatility = Q.vol.train$volatility

    reg.val = Q_features[window_ub,]
    reg.val$volatility = Q.vol.val$volatility
    
    # Fit model and predict
    lm.models[[i]]=lm(volatility ~ mean_BAS + var_order + var_BAS + var_volIm + var_bidSpread,
                      reg, weights = 0.8 ^ ((window_size - 1):0))
    r.squared.lm = c(r.squared.lm, summary(lm.models[[i]])$r.squared)
    pred.lm[[i]] = predict(lm.models[[i]], newdata = reg.val)
    
    # Accuracy measures
    MSE.lm = c(MSE.lm, mean((Q.vol.val$volatility - pred.lm[[i]]) ^ 2))
    QLIKE.lm = c(QLIKE.lm, mean(Q.vol.val$volatility / pred.lm[[i]] - 
                                   log(Q.vol.val$volatility / pred.lm[[i]]) - 1))
    MAE.lm = c(MAE.lm, abs(Q.vol.val$volatility - pred.lm[[i]]))
    RMSPE.lm = c(RMSPE.lm, mean(((Q.vol.val$volatility - pred.lm[[i]])/Q.vol.val$volatility)^2))
    # Increment window
    window_lb = window_lb + increment
    i = i + 1
  }
  return(list(lm.models, r.squared.lm, pred.lm, MSE.lm, QLIKE.lm, MAE.lm, RMSPE.lm))
}
```

```{r, warning=FALSE}
quarters.lm.models = quarters.pred.lm = list()
quarters.r.squared.lm = quarters.MSE.lm = quarters.QLIKE.lm = 
  quarters.MAE.lm = quarters.RMSPE.lm = vector()

for(q in 1:length(quarters)){
  # Fit linear models using rolling window
  result = fit_lm(quarters.vol[[q]], quarters.features[[q]])
  # Save results
  quarters.lm.models[[q]] = result[[1]]
  quarters.pred.lm[[q]] = result[[3]]
  quarters.r.squared.lm = c(quarters.r.squared.lm, result[[2]])
  quarters.MSE.lm = c(quarters.MSE.lm, result[[4]])
  quarters.QLIKE.lm = c(quarters.QLIKE.lm, result[[5]])
  quarters.MAE.lm =  c(quarters.MAE.lm, result[[6]])
  quarters.RMSPE.lm = c(quarters.RMSPE.lm, result[[7]])
}
```

### HAV-WLS

```{r}
fit_HAV = function(Q_vol, Q_quar){
  # Rolling window
  window_size = 12 # 2 hours = 12 buckets of 10-min
  window_lb = 1 
  increment = 1 # increment by 10 minutes each time
  max_index = nrow(Q_vol)
  i = 1
  
  HAV.wls.models = pred.HAV = list()
  r.squared.HAV = MSE.HAV = QLIKE.HAV = MAE.HAV = RMSPE.HAV = vector()
  
  while(max_index > window_lb + window_size - 1){
    window_ub = window_lb + (window_size - 1)
    # Use 2-hour data to train model (12 10-min time buckets)
    Q.vol.train = Q_vol[window_lb:window_ub,]
    # Actual volatility of the next 10 minutes (next time bucket)
    Q.vol.val = Q_vol[window_ub + 1,]
    # Realised quarticity 
    quar.train = Q_quar[window_lb:window_ub,]
    
    # Compute 5-day average volatility 
    mean.vol <- rep(0, window_size - 5)
    for (j in 1 : 5) {
      mean.vol <- mean.vol + Q.vol.train$volatility[j : (j + window_size - 6)] / 5
    }
    
    # Create dataframe for training & test sets
    HAV.df <- data.frame(vol = Q.vol.train$volatility[-(1:5)], 
                          vol_1 = Q.vol.train$volatility[5:(window_size - 1)],
                          mean_vol_5 = mean.vol)
    HAV.val <- data.frame(vol_1 = Q.vol.train$volatility[window_size],
                          mean_vol_5 = mean(Q.vol.train$volatility[(window_size-4): window_size]))
    
    # Fit model and predict
    HAV.wls.models[[i]] <- lm(vol ~ vol_1 + mean_vol_5, HAV.df,
                              weights = HAV.df$vol_1 / 
                                sqrt(quar.train$quarticity[5:(window_size - 1)]))
    r.squared.HAV = c(r.squared.HAV, summary(HAV.wls.models[[i]])$r.squared)
    pred.HAV[[i]] = predict(HAV.wls.models[[i]], newdata = HAV.val)
    
    # Accuracy measures
    MSE.HAV= c(MSE.HAV, mean((Q.vol.val$volatility - pred.HAV[[i]]) ^ 2))
    QLIKE.HAV = c(QLIKE.HAV, mean(Q.vol.val$volatility / pred.HAV[[i]] - 
                                   log(Q.vol.val$volatility / pred.HAV[[i]]) - 1))
    MAE.HAV= c(MAE.HAV, abs(Q.vol.val$volatility - pred.HAV[[i]]))
    RMSPE.HAV = c(RMSPE.HAV, mean(((Q.vol.val$volatility - pred.HAV[[i]])/
                                     Q.vol.val$volatility)^2))
    # Increment window
    window_lb = window_lb + increment
    i = i + 1
  }
  
  return(list(HAV.wls.models, r.squared.HAV, pred.HAV, MSE.HAV, QLIKE.HAV, MAE.HAV, RMSPE.HAV))
}
```

```{r}
quarters.HAV.models = quarters.pred.HAV = list()
quarters.r.squared.HAV = quarters.MSE.HAV = quarters.QLIKE.HAV = 
  quarters.MAE.HAV = quarters.RMSPE.HAV = vector()

for(q in 1:length(quarters)){
  # Fit HAV model using rolling window
  results = fit_HAV(quarters.vol[[q]], quarters.quarticity[[q]])
  # Save results
  quarters.HAV.models[[q]] = results[[1]]
  quarters.pred.HAV[[q]] = results[[3]]
  quarters.r.squared.HAV = c(quarters.r.squared.HAV, results[[2]])
  quarters.MSE.HAV = c(quarters.MSE.HAV, results[[4]])
  quarters.QLIKE.HAV = c(quarters.QLIKE.HAV, results[[5]])
  quarters.MAE.HAV =  c(quarters.MAE.HAV, results[[6]])
  quarters.RMSPE.HAV = c(quarters.RMSPE.HAV, results[[7]])
}
```

### Random forest

```{r}
fit_rf = function(Q_vol, Q_features){
  set.seed(16)
  
  # Rolling window
  window_size = 12 # 2 hours = 12 buckets of 10-min
  window_lb = 2 # start from 2nd time bucket
  increment = 1 # increment by 10 minutes each time
  max_index = nrow(Q_vol)
  i = 1
  
  rf.models = pred.rf = list()
  MSE.rf = QLIKE.rf = MAE.rf = RMSPE.rf = vector()
  
  while(max_index > window_lb + window_size - 1){
    window_ub = window_lb + (window_size - 1)
    # Use 2-hour data to train model (12 10-min time buckets)
    Q.vol.train = Q_vol[window_lb:window_ub,]
    # Actual volatility of the next 10 minutes (next time bucket)
    Q.vol.val = Q_vol[window_ub + 1,]
    
    idx = (window_lb - 1):(window_ub - 1)
    reg = Q_features[idx,]
    reg$volatility = Q.vol.train$volatility

    reg.val = Q_features[window_ub,]
    reg.val$volatility = Q.vol.val$volatility
    
    # Fit model and predict
    rf.models[[i]] = randomForest(volatility ~ mean_BAS + var_BAS + var_order 
                                  + var_volIm + var_bidSpread,
                                   reg, weights = 0.8 ^ ((window_size - 1):0))
    pred.rf[[i]] = predict(rf.models[[i]], newdata = reg.val)
    
    # Compute accuracy measures
    MSE.rf = c(MSE.rf, mean((Q.vol.val$volatility - pred.rf[[i]]) ^ 2))
    QLIKE.rf = c(QLIKE.rf, mean(Q.vol.val$volatility / pred.rf[[i]] - 
                                   log(Q.vol.val$volatility / pred.rf[[i]]) - 1))
    MAE.rf= c(MAE.rf, abs(Q.vol.val$volatility - pred.rf[[i]]))
    RMSPE.rf = c(RMSPE.rf, mean(((Q.vol.val$volatility - pred.rf[[i]])/Q.vol.val$volatility)^2))
    # Increment window
    window_lb <- window_lb + increment
    i <- i + 1
  }
  return(list(rf.models, pred.rf, MSE.rf, QLIKE.rf, MAE.rf, RMSPE.rf))
}
```

```{r}
quarters.rf.models = quarters.pred.rf = list()
quarters.MSE.rf = quarters.QLIKE.rf = quarters.MAE.rf = quarters.RMSPE.rf = vector()

for(q in 1:length(quarters)){
  # Fit random forest & predict using rolling window
  result = fit_rf(quarters.vol[[q]], quarters.features[[q]])
  # Save results
  quarters.rf.models[[q]] = result[[1]]
  quarters.pred.rf[[q]] = result[[2]]
  quarters.MSE.rf = c(quarters.MSE.rf, result[[3]])
  quarters.QLIKE.rf = c(quarters.QLIKE.rf, result[[4]])
  quarters.MAE.rf =  c(quarters.MAE.rf, result[[5]])
  quarters.RMSPE.rf = c(quarters.RMSPE.rf, result[[6]])
}
```

### SVM

```{r}
fit_SVM = function(Q_vol, Q_features){
  # Rolling window setup
  window_size <- 12  # 2 hours = 12 buckets of 10-min each
  window_lb <- 2     # Start from the second time bucket
  increment <- 1     # Increment by 10 minutes each time
  max_index <- nrow(Q_vol)
  i <- 1
  svm.models = pred.svm = list()
  MSE.svm = QLIKE.svm = MAE.svm = RMSPE.svm = vector()
  
  while (max_index > window_lb + window_size - 1) {
    window_ub <- window_lb + (window_size - 1)
    # Use 2-hour data to train the model (12 10-min time buckets)
    Q.vol.train <- Q_vol[window_lb:window_ub, ]
    # Actual volatility of the next 10 minutes (next time bucket)
    Q.vol.val <- Q_vol[window_ub + 1, ]
    
    idx = (window_lb - 1):(window_ub - 1)
    reg = Q_features[idx,]
    reg$volatility = Q.vol.train$volatility

    reg.val = Q_features[window_ub,]
    reg.val$volatility = Q.vol.val$volatility
    
    # Train SVM model
    svm.models[[i]] <- svm(
      volatility ~ mean_BAS + var_BAS + var_order + var_volIm + var_bidSpread, 
      data = reg,
      kernel = "radial",  
      cost = 1            
    )
    # Predict the next value using the SVM model
    pred.svm[[i]] <- predict(svm.models[[i]], newdata = reg.val)
    
    # Calculate accuracy measures
    MSE.svm <- c(MSE.svm, mean((Q.vol.val$volatility - pred.svm[[i]]) ^ 2))
    QLIKE.svm <- c(QLIKE.svm, mean(Q.vol.val$volatility / pred.svm[[i]] - 
                                  log(Q.vol.val$volatility / pred.svm[[i]]) - 1))
    MAE.svm = c(MAE.svm, abs(Q.vol.val$volatility - pred.svm[[i]]))
    RMSPE.svm = c(RMSPE.svm, mean(((Q.vol.val$volatility - pred.svm[[i]])/
                                     Q.vol.val$volatility)^2))
    # Increment the window
    window_lb <- window_lb + increment
    i = i + 1
  }
  return(list(svm.models, pred.svm, MSE.svm, QLIKE.svm, MAE.svm, RMSPE.svm))
}
```

```{r}
# Initialize variables to store results
quarters.svm.models = quarters.pred.svm = list()  # Store all quarter SVM models
quarters.MSE.svm = quarters.QLIKE.svm = quarters.MAE.svm = quarters.RMSPE.svm = vector() 

# Loop through each group of quarters (replace `quarters` with your list of time ID groupings)
for (q in 1:length(quarters)) {
  # Fit SVM & predict using rolling window
  results = fit_SVM(quarters.vol[[q]], quarters.features[[q]])
  # Save results
  quarters.svm.models[[q]] = results[[1]]
  quarters.pred.svm[[q]] = results[[2]]
  quarters.MSE.svm =  c(quarters.MSE.svm, results[[3]])
  quarters.QLIKE.svm = c(quarters.QLIKE.svm, results[[4]])
  quarters.MAE.svm =  c(quarters.MAE.svm, results[[5]])
  quarters.RMSPE.svm = c(quarters.RMSPE.svm, results[[6]])
}
```

## Accuracy

### Linear Regression

```{r}
boxplot(quarters.r.squared.lm, horizontal = TRUE, main='Distribution of R2')
boxplot(quarters.MSE.lm, horizontal = TRUE, main='Distribution of MSE')
boxplot(quarters.QLIKE.lm, horizontal = TRUE, main='Distribution of QLIKE')
boxplot(quarters.MAE.lm, horizontal = TRUE, main='Distribution of MAE')
boxplot(quarters.RMSPE.lm, horizontal = TRUE, main='Distribution of RMSPE')

acc.lm = data.frame(MSE = quarters.MSE.lm,
                    QLIKE = quarters.QLIKE.lm,
                    MAE = quarters.MAE.lm,
                    RMSPE = quarters.RMSPE.lm)
save(acc.lm, file='../data/lm_acc.RDS')
```

### HAV-WLS

```{r}
boxplot(quarters.r.squared.HAV, horizontal = TRUE, main='Distribution of R2')
boxplot(quarters.MSE.HAV , horizontal = TRUE, main='Distribution of MSE')
boxplot(quarters.QLIKE.HAV, horizontal = TRUE, main='Distribution of QLIKE')
boxplot(quarters.MAE.HAV, horizontal = TRUE, main='Distribution of MAE')
boxplot(quarters.RMSPE.HAV, horizontal = TRUE, main='Distribution of RMSPE')

acc.HAV = data.frame(MSE = quarters.MSE.HAV,
                    QLIKE = quarters.QLIKE.HAV,
                    MAE = quarters.MAE.HAV,
                    RMSPE = quarters.RMSPE.HAV)
save(acc.HAV, file='../data/HAV_acc.RDS')
```

### Random forest

```{r}
boxplot(quarters.MSE.rf , horizontal = TRUE, main='Distribution of MSE')
boxplot(quarters.QLIKE.rf , horizontal = TRUE, main='Distribution of QLIKE')
boxplot(quarters.MAE.rf , horizontal = TRUE, main='Distribution of MAE')
boxplot(quarters.RMSPE.rf , horizontal = TRUE, main='Distribution of RMSPE')

acc.rf = data.frame(MSE = quarters.MSE.rf,
                    QLIKE = quarters.QLIKE.rf,
                    MAE = quarters.MAE.rf,
                    RMSPE = quarters.RMSPE.rf)
save(acc.rf, file='../data/rf_acc.RDS')
```

### SVM

```{r}
boxplot(quarters.MSE.svm , horizontal = TRUE, main='Distribution of MSE')
boxplot(quarters.QLIKE.svm , horizontal = TRUE, main='Distribution of QLIKE')
boxplot(quarters.MAE.svm , horizontal = TRUE, main='Distribution of MAE')
boxplot(quarters.RMSPE.svm , horizontal = TRUE, main='Distribution of RMSPE')

acc.svm = data.frame(MSE = quarters.MSE.svm,
                    QLIKE = quarters.QLIKE.svm,
                    MAE = quarters.MAE.svm,
                    RMSPE = quarters.RMSPE.svm)
save(acc.svm, file='../data/svm_acc.RDS')
```

### Summary

```{r}
# Median prediction errors
median_acc = data.frame(colMedians(acc.lm, na.rm=TRUE),
                        colMedians(acc.HAV, na.rm=TRUE),
                        colMedians(acc.rf, na.rm=TRUE),
                        colMedians(acc.svm, na.rm=TRUE))
colnames(median_acc) = c('LR', 'HAV', 'RF', 'SVM')
median_acc
```

```{r}
# Accuracy ranking
acc_rank = apply(median_acc, 1, function(x) rank(x, ties.method = "min"))
acc_rank = as.data.frame(t(acc_rank))
acc_rank
```

```{r}
# Compute accuracy score
error_normalised = as.data.frame(lapply(median_acc, 
                                        function(x) x/rowSums(median_acc)))
acc_score = as.data.frame(t(1/colMeans(error_normalised)))
colnames(acc_score) = c('Linear Regression', 'HAV-WLS', 
                         'Random Forest', 'Support Vector Machine')
acc_score
```

## Robustness

```{r, warning=FALSE}
# Function to replace a percentage of volatility values with random values
replace_data <- function(data, percentage, lower = 0.004, upper = 0.008) {
  set.seed(2024)
  n_replace <- floor(nrow(data) * percentage)  # Number of elements to replace
  idx <- sample(nrow(data), size = n_replace, replace = FALSE)
  
  # Replace the data at those indices with random values between lower and upper
  df = data
  for (i in 1:length(idx)) {
    df[idx[i], 2] <- runif(1, min = lower, max = upper)
  }
  return(df)
}

# Function to fit models on contaminated data and make predictions
volatility_prediction_with_random_replacement <- function(percentage, model) {
  replaced_pred = list()
  for (i in 1:length(quarters)){
    replaced_vol = replace_data(quarters.vol[[i]], percentage)
    if(model == 'lm'){
      pred = fit_lm(replaced_vol, quarters.features[[i]])[[3]]
    } else if(model == 'hav'){
      pred = fit_HAV(replaced_vol, quarters.quarticity[[i]])[[3]]
    } else if(model == 'rf'){
      pred = fit_rf(replaced_vol, quarters.features[[i]])[[2]]
    } else if(model == 'svm'){
      pred = fit_SVM(replaced_vol, quarters.features[[i]])[[2]]
    }
    replaced_pred[[i]] = pred
  }
  return(replaced_pred)
}

# Function to calculate the area between two lines using the trapezoidal rule
calculate_area <- function(line1, line2) {
  abs_diff <- abs(line1 - line2)
  area <- trapz(abs_diff) 
  return(area)
}
```

```{r, warning=FALSE}
models = c('lm', 'hav', 'rf', 'svm')
models.pred = list(quarters.pred.lm, quarters.pred.HAV, quarters.pred.rf, quarters.pred.svm)
robust.lists = robust.plots = robust.df = robust.areas = list()
for(i in 1:length(models)){
  train_5_pred = volatility_prediction_with_random_replacement(0.05, models[i])
  train_10_pred = volatility_prediction_with_random_replacement(0.10, models[i])
  train_15_pred = volatility_prediction_with_random_replacement(0.15, models[i])
  quarters.pred = models.pred[[i]]
  # Averaging the 4 different predictions of the quarters
  train_5_pred <- mapply(mean, train_5_pred[[1]], train_5_pred[[2]], 
                        train_5_pred[[3]], train_5_pred[[4]], na.rm = TRUE)
  train_10_pred <- mapply(mean, train_10_pred[[1]], train_10_pred[[2]], 
                          train_10_pred[[3]], train_10_pred[[4]], na.rm = TRUE)
  train_15_pred <- mapply(mean, train_15_pred[[1]], train_15_pred[[2]], 
                          train_15_pred[[3]], train_15_pred[[4]], na.rm = TRUE)
  train_0_pred <- mapply(mean, quarters.pred[[1]], quarters.pred[[2]], 
                      quarters.pred[[3]], quarters.pred[[4]], na.rm = TRUE)
  # Create dataframe for shiny
  robust.df[[i]] = data.frame(time = 1:length(train_0_pred), pred = train_0_pred, 
                              train_5 = train_5_pred, train_10 = train_10_pred, 
                              train_15 = train_15_pred)
  
  df = lapply(list(pred = train_0_pred, train_5 = train_5_pred, 
                    train_10 = train_10_pred, train_15 = train_15_pred), 
               function(x) data.frame(Time = 1:length(x), Volatility = unlist(x)))
  combined.df = do.call(rbind, lapply(names(df), 
                                    function(x) {cbind(df[[x]], Condition = x)}))
  combined.df$Condition = factor(combined.df$Condition, 
                                 levels=c('pred', 'train_5', 'train_10', 'train_15'))
  
  # Visualise adversarial areas
  robust.plots[[i]] = ggplot(combined.df, 
                             aes(x = Time, y = Volatility, colour = Condition)) +
                     geom_line() +
                     geom_point(size = 1, alpha = 0.6) +
                     labs(
                      title = "Predicted Volatility using Contaminated Training Data",
                      x = "Time (index)",
                      y = "Predicted Volatility",
                      colour = "Condition") +
                    theme_minimal() 
  
  # Extracting the relevant data from the combined data frame
  true_pred <- combined.df$Volatility[combined.df$Condition == 'pred']
  train_5 <-  combined.df$Volatility[combined.df$Condition == 'train_5']
  train_10 <- combined.df$Volatility[combined.df$Condition == 'train_10']
  train_15 <- combined.df$Volatility[combined.df$Condition == 'train_15']
  # Calculating the areas
  area.train_5 <- calculate_area(true_pred, train_5)
  area.train_10 <- calculate_area(true_pred, train_10)
  area.train_15 <- calculate_area(true_pred, train_15)
  # Tabulating the results
  robust.areas[[i]] = data.frame(
    `Comparison` = c("0% vs 5%", "0% vs 10%", "0% vs 15%"),
    `Area` = c(area.train_5, area.train_10, area.train_15)
  )
}
```

```{r}
#Save data
save(robust.df, file='../data/robust_vol.RDS')
save(robust.areas, file = '../data/robust_area.RDS')
```

### Linear regression

```{r}
robust.plots[[1]]
robust.areas[[1]]
```

### HAV

```{r}
robust.plots[[2]]
robust.areas[[2]]
```

### Random forest

```{r}
robust.plots[[3]]
robust.areas[[3]]
```

### SVM

```{r}
robust.plots[[4]]
robust.areas[[4]]
```

### Summary

```{r}
names(robust.areas[[1]]) <- c("Comparison", "Area LR")
names(robust.areas[[2]]) <- c("Comparison", "Area HAV")
names(robust.areas[[3]]) <- c("Comparison", "Area RF")
names(robust.areas[[4]]) <- c("Comparison", "Area SVM")
combined_area <- Reduce(function(x, y) merge(x, y, by = "Comparison", all = TRUE), robust.areas)
comparison_order <- factor(combined_area$Comparison, levels = c("0% vs 5%", "0% vs 10%", "0% vs 15%"))
combined_area <- combined_area[order(comparison_order), ]
combined_area
save(combined_area, file='../data/robust_area.Rds')
```

```{r}
# Robustness ranking
combined_area$rank_LR <- apply(combined_area[, -1], 1, function(x) rank(x, ties.method = "min")[1])
combined_area$rank_HAV <- apply(combined_area[, -1], 1, function(x) rank(x, ties.method = "min")[2])
combined_area$rank_RF <- apply(combined_area[, -1], 1, function(x) rank(x, ties.method = "min")[3])
combined_area$rank_SVM <- apply(combined_area[, -1], 1, function(x) rank(x, ties.method = "min")[4])
area_ranking_table <- combined_area[, c("Comparison", "rank_LR", "rank_HAV", "rank_RF", "rank_SVM")]
comparison_order <- factor(area_ranking_table$Comparison, levels = c("0% vs 5%", "0% vs 10%", "0% vs 15%"))
area_ranking_table <- area_ranking_table[order(comparison_order), ]
colnames(area_ranking_table) = c("Comparison", "Rank LR", "Rank HAV", "Rank RF", "Rank SVM")
area_ranking_table

save(area_ranking_table, file='../data/area_ranking.RDS')
```

```{r}
# Robustness scores
data_to_normalise = combined_area[, 2:5]
combined_area_normalised = as.data.frame(lapply(data_to_normalise, 
                                                function(x) x / rowSums(data_to_normalise)))
robustness.scores_per_area = colSums(combined_area_normalised)
robust_score = as.data.frame(t(1/robustness.scores_per_area))
colnames(robust_score) = c('LR', 'HAV', 'RF', 'SVM')
robust_score
```

## Interpretability

```{r}
survey = read.csv("../data/cleaned_survey.csv")
survey
```

```{r, message=FALSE}
cleaned_survey = survey

# Process survey data for visualisation
survey$Q1[survey$Q1 == 'Zero/minimal (No/very little exposure to financial concepts/terminology)'] <- 'Zero/minimal'
survey$Q1[survey$Q1 == 'Intermediate proficiency (Some understanding of financial concepts/terminology)'] <- 'Intermediate'
survey$Q1[survey$Q1 == 'Advanced (Fluency in financial concepts/terminology)'] <- 'Advanced'

survey$Q2[survey$Q2 == 'Zero/minimal (No/very little exposure to machine learning concepts/terminology)'] <- 'Zero/minimal'
survey$Q2[survey$Q2 == 'Intermediate proficiency (Some understanding of machine learning concepts/terminology)'] <- 'Intermediate'
survey$Q2[survey$Q2 == 'Advanced (Fluency in machine learning concepts/terminology)'] <- 'Advanced'

survey$Q3[survey$Q3== 0] <- "No understanding"
survey$Q3[survey$Q3== 0.25] <- "Limited understanding"
survey$Q3[survey$Q3== 0.5] <- "Basic understanding"
survey$Q3[survey$Q3== 0.75] <- "Good understanding"
survey$Q3[survey$Q3== 1] <- "Strong understanding"
survey$Q3 = factor(survey$Q3, levels=c('No understanding', 'Limited understanding', 'Basic understanding', 'Good understanding', 'Strong understanding'))

survey$Q6[survey$Q6== 0] <- "No understanding"
survey$Q6[survey$Q6== 0.25] <- "Limited understanding"
survey$Q6[survey$Q6== 0.5] <- "Basic understanding"
survey$Q6[survey$Q6== 0.75] <- "Good understanding"
survey$Q6[survey$Q6== 1] <- "Strong understanding"
survey$Q6 = factor(survey$Q6, levels=c('No understanding', 'Limited understanding', 'Basic understanding', 'Good understanding', 'Strong understanding'))

survey$Q9[survey$Q9== 0] <- "No understanding"
survey$Q9[survey$Q9== 0.25] <- "Limited understanding"
survey$Q9[survey$Q9== 0.5] <- "Basic understanding"
survey$Q9[survey$Q9== 0.75] <- "Good understanding"
survey$Q9[survey$Q9== 1] <- "Strong understanding"
survey$Q9 = factor(survey$Q9, levels=c('No understanding', 'Limited understanding', 'Basic understanding', 'Good understanding', 'Strong understanding'))

survey$Q14[survey$Q14== 0] <- "No understanding"
survey$Q14[survey$Q14== 0.25] <- "Limited understanding"
survey$Q14[survey$Q14== 0.5] <- "Basic understanding"
survey$Q14[survey$Q14== 0.75] <- "Good understanding"
survey$Q14[survey$Q14== 1] <- "Strong understanding"
survey$Q14 = factor(survey$Q14, levels=c('No understanding', 'Limited understanding', 'Basic understanding', 'Good understanding', 'Strong understanding'))

survey$LinearTest = factor(survey$LinearTest, levels=c(0, 1, 2))
survey$HAVTest = factor(survey$HAVTest, levels=c(0, 1, 2))
survey$RFTest = factor(survey$RFTest, levels=c(0, 1, 2))
survey$SVMTest = factor(survey$SVMTest, levels=c(0, 1))

save(survey, file='../data/survey.RData')
```

```{r, message=FALSE}
# Visualisation of survey results
p1 = survey |>
    group_by(Q3, Q2) |> 
    summarise(count = n()) |> 
    ungroup() |> 
    ggplot(aes(x=Q3, y=count, fill=Q2)) +
      geom_bar(stat="identity", position = 'stack', alpha = 0.75) + 
      labs(x='Understanding Confidence', y='Count', fill='Machine learning proficiency') + 
      theme(legend.position = "bottom")
p1

p2 = survey |>
    group_by(Q3, Q1) |> 
    summarise(count = n()) |> 
    ungroup() |> 
    ggplot(aes(x=Q3, y=count, fill=Q1)) +
      geom_bar(stat="identity", position = 'stack', alpha = 0.75) + 
      labs(x='Understanding Confidence', y='Count', fill='Finance proficiency') +
      theme(legend.position = "bottom")
p2

p3 = survey |>
    group_by(LinearTest, Q1) |> 
    summarise(count = n()) |> 
    ungroup() |> 
    ggplot(aes(x=LinearTest, y=count, fill=Q1)) +
      geom_bar(stat="identity", position = 'stack', alpha = 0.75) + 
      labs(x='Number of correct answers', y='Count', fill='Finance proficiency')+
      theme(legend.position = "bottom")
p3

p4 = survey |>
    group_by(LinearTest, Q2) |> 
    summarise(count = n()) |> 
    ungroup() |> 
    ggplot(aes(x=LinearTest, y=count, fill=Q2)) +
      geom_bar(stat="identity", position = 'stack', alpha = 0.75) + 
      labs(x='Number of correct answers', y='Count', fill='Machine learning proficiency') +
      theme(legend.position = "bottom")
p4
```

```{r}
technical <- cleaned_survey |> filter(
  Q1 == "Intermediate proficiency (Some understanding of financial concepts/terminology)" |
    Q1 == "Advanced (Fluency in financial concepts/terminology)")

general <- cleaned_survey |> filter(
  Q1 == "Zero/minimal (No/very little exposure to financial concepts/terminology)")

# Obtain interpretability scores for 2 stakeholder groups
inter_score = data.frame(c('Technical', 'General'),
                         c(mean(technical$LinearTechnical), mean(general$LinearGeneral)),
                         c(mean(technical$HAVTechnical), mean(general$HAVGeneral)),
                         c(mean(technical$RFTechnical), mean(general$RFGeneral)),
                         c(mean(technical$SVMTechnical), mean(general$SVMGeneral)))

colnames(inter_score) = c('Stakeholder', 'Linear Regression', 'HAV-WLS', 
                          'Random Forest', 'Support Vector Machine')
inter_score
save(inter_score, file = '../data/inter_score.RData')
```

## Model selection

```{r}
acc_score = as.numeric(acc_score)
robust_score = as.numeric(robust_score)
tech_inter_score = as.numeric(inter_score[1,2:5])
general_inter_score = as.numeric(inter_score[2,2:5])

# Min-max normalisation of performance metric scores
norm_acc_score = (acc_score - min(acc_score))/(max(acc_score)-min(acc_score))
norm_robust_score = (robust_score - min(robust_score))/
                        (max(robust_score)-min(robust_score))
norm_tech_inter_score = (tech_inter_score - min(tech_inter_score))/
                        (max(tech_inter_score)- min(tech_inter_score))
norm_general_inter_score = (general_inter_score - min(general_inter_score))/
                        (max(general_inter_score)- min(general_inter_score))
```

```{r}
# Summary of normalised scores of performance metrics
eval = data.frame('Model' = c('LR', 'HAV', 'RF', 'SVM'),
                  'Accuracy' = norm_acc_score,
                  'Robustness' = norm_robust_score,
                  'Technical Interpretability' = norm_tech_inter_score,
                  'General Interpretability' = norm_general_inter_score)
eval
```

```{r}
# Plot performance metrics against each other
ggplot(eval, aes(x=Accuracy, y=Robustness, color=Model)) +
  geom_point(size = 3) +
  theme_minimal() +
  labs(x='Interpretability', y ='Accuracy')
```

```{r}
# Compute weighted average score
## Technical - accuracy > robustness > interpretability
tech_model_score = 0.5*norm_acc_score + 0.3*norm_robust_score + 0.2*norm_tech_inter_score
tech_model_score
## General - interpretability > accuracy > robustness
general_model_score = 0.25*norm_acc_score + 0.1*norm_robust_score + 0.65*norm_general_inter_score
general_model_score
```
